\section{Related Work\label{sec:topologies.related}}

The motivation behind topology generation originates from the need to evaluate network protocols in simulations.
In fact, while network topology should not influence the behavior of a protocol, it can significantly impact its performance~\cite{tangmunarunkit_Networktopologygenerators_2002}.
Multiple tools have been developed to generate network topologies at the time, such as \texttt{GT-ITM}~\cite{calvert_ModelingInternettopology_1997}, \texttt{Tiers}~\cite{doar_bettermodelgenerating_1996}, or \texttt{BRITE}~\cite{medina_BRITEapproachuniversal_2001}.
\Textcite{tangmunarunkit_Networktopologygenerators_2002} distinguish two main categories of topology generators: \emph{structural}, which aim at reproducing the structural properties of the internet and particularly its hierarchical organization, and \emph{degree-based}, which focus on the statistical properties of the network, notably the power-law distribution of the node degrees~\cite{faloutsos_powerlawrelationshipsInternet_1999}.
Most of these works are more than 20 years old and have been developed to generate topologies for internet-scale networks, which are not directly applicable to the generation of \glspl{fids} datasets, particularly in our siloed IT networks.

Recent works on topology generation are rarer and focus on specific use cases.
For instance, \textcite{laurito_TopoGennetworktopology_2017} developed \texttt{TopoGen}, a tool to generate network topologies using \glspl{sdn}.
Their approach allows users to programmatically define the network topology using the Ruby programming language, and extract existing topologies from real-world networks.
Yet, their approach is limited to \glspl{sdn} and does not allow automating data generation.
\Textcite{alrumaih_GENINDindustrialnetwork_2023} developed \texttt{GENIND}, a tool to generate industrial network topologies.
Similarly to us, the authors identified that most existing tools are too focused on internet-inspired and internet-scale topologies, and do not allow generating topologies for specific use cases.
Their tool focus on generating topologies for industrial networks, and therefore generates topologies with specific constraints layer-by-layer, before connecting the different sub-topologies together in a multigraph.
To the best of our knowledge, no existing tool allows generating network topologies for IT networks, that can be randomized to obtain variations with common characteristics.

Another aspect of this work is building topologies that satisfy the requirements of specified attack scenarios, and deploying them in a controlled environment.
Recent works have focused on cyber ranges\footnotemark{} for the latter.
Both \textcite{venkatesan_VulnerVANVulnerableNetwork_2019,yamin_Modelingexecutingcyber_2022} translate a scenario and a topology description into a deployable environment, but do not focus on the generation of the topologies themselves.
More recently, \textcite{besson_URSIDAutomaticallyRefining_2024a} developed \texttt{URSID}, a tool to generate vulnerable network architectures that implement user provided attack scenarios.
Compared to previous works, \texttt{URSID} introduces variability in the scenarios and service deployments, but does not provide heterogeneous topologies constructions.
Therefore, no existing works provide a complete solution to generate heterogeneous \gls{fids} datasets.

\footnotetext{A cyber range is a controlled environment simulating a real-world network, where users can deploy their own topologies and scenarios, offering a wide range of applications, from cybersecurity training to dataset generation~\cite{nationalinitiativeforcybersecurityeducation_CyberRangeGuide_2023}.}